### 2018-03-14 paper/2-background.tex
Work on S3
Polish S2

When build systems are used by large teams, different team members often end up
executing exactly the same tasks on their local machines. A \emph{cloud build
system} can speed up builds dramatically by transparently sharing build results
among team members. Furthermore, cloud build systems allow one to perform
\emph{shallow builds} that materialise only end build products locally, leaving
all intermediates in the cloud.

\Bazel is one of the first examples of openly available cloud build systems.
Like \Make, it does not support dynamic dependencies and can therefore benefit
from the simplicity of building tasks in a statically known topological order.
It is minimal and supports the early cutoff optimisation.

To support cloud builds, \Bazel maintains a \emph{content-addressable cache},
which can be used to fetch a previously built file given the hash of its
content, as well as \emph{dependency graphs from all previous builds}, annotated
with observed file hashes\footnote{Here we ignore the issue of limited cloud
storage resources for the sake of simplicity; in practice, old entries are
regularly evicted from the storage, as further discussed
in~\S\ref{sec-cloud-aspects}.}. The latter allows to bypass the execution of
a task, by predicting the hash of the result from the hashes of its dependencies,
and subsequently fetching the result from the cache. A concrete implementation is
provided in~\S\ref{sec-implementations}.

Consider a cloud build example in Fig.~\ref{fig-bazel}. The user starts by
checking out the project sources, whose hashes happen to be 1, 2 and 3, and
requests to build \cmd{main.exe}. By looking up the history of all previous
builds, the build system finds that someone has already compiled these sources
before and obtains the hashes (4~and~5) of the resulting \cmd{util.o} and
\cmd{main.o} files (the files themselves do not need to be downloaded).
Similarly, the build system finds that the hash of the resulting \cmd{main.exe}
should be 6 and downloads the actual binary from the cache, since it is the end
build product and must be materialised. In the second scenario, shown in
Fig.~\ref{fig-bazel}(c), the user modifies \cmd{util.c}, thereby changing its
hash from 1 to 7. The build system finds that nobody has ever compiled the new
$\{\cmd{util.c}, \cmd{util.h}\}$ combination and must therefore build
\cmd{util.o}, which results in changing its hash from~4~to~8. The obtained
combination of hashes of \cmd{util.o} and \cmd{main.o} has not been encountered
before, therefore the build system first downloads \cmd{main.o} from the cache
and then builds \cmd{main.exe} by linking the two object files. When the build
is complete, the obtained files as well as the new dependency graph annotated
with hashes are stored in the cloud for future reuse.

@@

\begin{table}[h]
\smaller
\centering
\begin{tabular}{l||l|l||l|c|c|c}
\hline
$\!$Build system$\!$& Metadata                       & Algorithm    & Dependencies & Minimal & Cutoff & Cloud$\!$\\\hline
$\!$\Make       $\!$& File modification times        & Topological  & Static       & Yes     & No     & No   $\!$\\
$\!$\Excel      $\!$& Dirty cells, calculation chain & Reordering   & Dynamic      & No      & No     & No   $\!$\\
$\!$\Shake      $\!$& Previous dependency graph      & Recursive    & Dynamic      & Yes     & Yes    & No   $\!$\\
$\!$\Bazel      $\!$& All dependency graphs, cache   & Topological  & Static       & Yes     & Yes    & Yes  $\!$\\\hline
\hline
\end{tabular}
\vspace{2mm}
\caption{Summary of build system differences.\label{tab-summary}}
\vspace{-7mm}
\end{table}

>>>>

When build systems are used by large teams, different team members often end up
executing exactly the same tasks on their local machines. A \emph{cloud build
system} can speed up builds dramatically by transparently sharing build results
among team members. Furthermore, cloud build systems allow one to perform
\emph{shallow builds} that materialise only end build products locally, leaving
all intermediates in the cloud. We illustrate shallow cloud builds by an example
(see~Fig.~\ref{fig-bazel}).

The user starts by checking out the project sources, whose hashes are (for
simplicity)~1,~2~and~3, and requests to build \cmd{main.exe}, see
Fig.~\ref{fig-bazel}(a,b). By looking up the history of all previous builds of
the project\footnote{Here we ignore the issue of limited cloud storage resources
for the sake of simplicity; in practice, old entries are regularly evicted from
the storage, as further discussed in~\S\ref{sec-cloud-aspects}.}, the build
system finds that someone has already compiled these exact sources before and
their resulting files \cmd{util.o} and \cmd{main.o} had hashes 4~and~5.
Similarly, the build system finds that the hash of the resulting \cmd{main.exe}
should be 6 and downloads the actual binary from the shared cloud storage, since
it is the end build product and must therefore be materialised.

In the second scenario, shown in Fig.~\ref{fig-bazel}(c), the user modifies the
source \cmd{util.c}, thereby changing its hash from~1~to~7. The build system
finds that nobody has ever compiled the new $\{\cmd{util.c}, \cmd{util.h}\}$
combination and must therefore build \cmd{util.o}, which results in changing its
hash from~4~to~8. The obtained combination of hashes of \cmd{util.o} and
\cmd{main.o} has not been encountered before either, therefore the build system
first downloads \cmd{main.o} from the cloud and then builds \cmd{main.exe} by
linking the two object files. When the build is complete, the obtained results
may be uploaded to the cloud for future reuse by other team members.

\Bazel is one of the first examples of openly available cloud build systems.
Like \Make, it does not support dynamic dependencies and can therefore benefit
from the simplicity of building tasks in a statically known topological order.
It is minimal and supports the early cutoff optimisation. To support cloud
builds, \Bazel maintains a \emph{content-addressable cache} that can be used to
fetch a previously built file given the hash of its content, and
\emph{dependency graphs from all previous builds}, annotated with observed file
hashes. The latter allows to bypass the execution of a task, by predicting the
hash of the result from the hashes of its dependencies, and subsequently
fetch the result from the cache. A concrete implementation is provided
in~\S\ref{sec-implementations}.

@@

\begin{table}[h]
\vspace{-1mm}
\smaller
\centering
\begin{tabular}{l||l|l||l|c|c|c}
\hline
$\!$Build system$\!$& Metadata                       & Algorithm    & Dependencies & Minimal & Cutoff & Cloud$\!$\\\hline
$\!$\Make       $\!$& File modification times        & Topological  & Static       & Yes     & No     & No   $\!$\\
$\!$\Excel      $\!$& Dirty cells, calculation chain & Reordering   & Dynamic      & No      & No     & No   $\!$\\
$\!$\Shake      $\!$& Previous dependency graph      & Recursive    & Dynamic      & Yes     & Yes    & No   $\!$\\
$\!$\Bazel      $\!$& All dependency graphs, cache   & Topological  & Static       & Yes     & Yes    & Yes  $\!$\\\hline
\hline
\end{tabular}
\vspace{0.5mm}
\caption{Summary of build system differences.\label{tab-summary}}
\vspace{-8mm}
\end{table}
